Design and Implementation of Probabilistic Programming Language Anglican

### 2016-11-20 paper/paper.tex
removed typo

Checkpoints can be uniquely named at the compilation time;
however, at runtime a checkpoint corresponding to a single code
location may occur multiple time due to recurrent invocation of
the function containing the checkpoint. Every unique occurrence
of a checkpoint must receive a different address. An addressing
scheme based on computing of stack addresses of checkpoints was
described in the context of Lightweight
Metropolis-Hastings~\cite{WSG11}.  This scheme has advantages
over the naive scheme where dynamic occurrences of checkpoints
are numbered sequentially, however it impacts the 
performance of probabilistic programs because of the computation
cost of computing the stack addresses:
\begin{enumerate}
    \item On every function call, a component is added to the
        address, hence the address size is linear in stack depth.
    \item Every function call must be augmented by symbolic
        information required to compute stack addresses.
\end{enumerate}
In addition, the above scheme can still lead to inferior
correspondence between checkpoints in different traces: in
Anglican and other probabilistic languages where distributions
are first-class objects checkpoints with incompatible arguments 
can correspond to the same stack address. Consider the following
program fragment:
\begin{lstlisting}[style=default]
(let [dist (if use-gamma
               (gamma 2. 2.)
               (normal 0. 1.))
      x (sample dist)]
  (predict x))
\end{lstlisting}
The \texttt{sample} checkpoint has the same stack address in
different traces, however the random values should not be
re-used between different distributions. Further on, in some
algorithms, such as black-box variational inference, the role of 
checkpoint addresses is semantic rather than heuristic ---
appropriate correspondence must be established
between checkpoints in different traces for the algorithm to
work.

>>>>

Checkpoints can be uniquely named at compilation time;
however, at runtime a checkpoint corresponding to a single code
location may occur multiple time due to recurrent invocation of
the function containing the checkpoint. Every unique occurrence
of a checkpoint must receive a different address. An addressing
scheme based on computing of stack addresses of checkpoints was
described in the context of Lightweight
Metropolis-Hastings~\cite{WSG11}.  This scheme has advantages
over the naive scheme where dynamic occurrences of checkpoints
are numbered sequentially, however it impacts the 
performance of probabilistic programs because of the computation
cost of computing the stack addresses:
\begin{enumerate}
    \item On every function call, a component is added to the
        address, hence the address size is linear in stack depth.
    \item Every function call must be augmented by symbolic
        information required to compute stack addresses.
\end{enumerate}
In addition, the above scheme can still lead to inferior
correspondence between checkpoints in different traces: in
Anglican and other probabilistic languages where distributions
are first-class objects checkpoints with incompatible arguments 
can correspond to the same stack address. Consider the following
program fragment:
\begin{lstlisting}[style=default]
(let [dist (if use-gamma
               (gamma 2. 2.)
               (normal 0. 1.))
      x (sample dist)]
  (predict x))
\end{lstlisting}
The \texttt{sample} checkpoint has the same stack address in
different traces, however the random values should not be
re-used between different distributions. Further on, in some
algorithms, such as black-box variational inference, the role of 
checkpoint addresses is semantic rather than heuristic ---
appropriate correspondence must be established
between checkpoints in different traces for the algorithm to
work.
